Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Simplify appendix on exact energy
Review appendix B

\appendix
\section{Computing the exact energy of the non-interacting system}
    Here we show how to arrive at \autoref{eq:exact_energy}. We need to solve
    two integrals consisting of $N$, $d$-dimensional, Gaussians from
    \autoref{eq:trial_simple_gaussian}. We will label the position in the $j$-th
    coordinate for the $i$-th particle by $x_{ij}$. Starting with the
    normalization factor we get
    \begin{align}
        \bra{\Psi_T}\ket{\Psi_T}
        &=
        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2
        \\
        &=
        \prod_{i = 1}^N
        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)
        \\
        &=
        \prod_{i = 1}^N
        \prod_{j = 1}^d
        \int\dd x_{ij}\exp(-2\alpha x_{ij}^2)
        \\
        &=
        \prod_{i = 1}^N
        \prod_{j = 1}^d
        \sqrt{\frac{\pi}{2\alpha}}
        = \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{dN}.
    \end{align}
    From the Hamiltonian we get two integrals containing $\vf{r}_i^2$ for each
    particle $i$.
    \begin{align}
        \bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}
        &=
        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2\vf{r}_i^2
        \\
        &=
        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)\vf{r}_i^2
        \nonumber \\
        &\qquad
        \times
        \prod_{j \neq i}^N
        \int\dd\vf{r}_j\exp(-2\alpha\vf{r}_j^2)
        \\
        &=
        \sum_{l = 1}^d
        \int
        \para{
            \prod_{k = 1}^d
            \dd x_{ik}\exp(-2\alpha x_{ik}^2)
        } x_{il}^2
        \nonumber\\
        &\qquad
        \times
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d(N - 1)}
        \\
        &=
        \sum_{l = 1}^d
        \frac{1}{4\alpha}\sqrt{\frac{\pi}{2\alpha}}
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d - 1}
        \nonumber\\
        &\qquad
        \times
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d(N - 1)}
        \\
        &= \frac{d}{4\alpha}\para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{dN}.
    \end{align}
    We thus have that
    \begin{align}
        \frac{\bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
        &=
        \frac{d}{4\alpha}.
    \end{align}
    Using \autoref{eq:laplacian_simple_gaussian} we can find the exact energy
    expression we are looking for.
    \begin{align}
        E(\alpha)
        &=
        \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
        \\
        &=
        \sum_{i = 1}^N
        \brac{
            -\frac{\hbar^2}{2m}
            \brak{
                -2d\alpha
                + 4\alpha^2\frac{d}{4\alpha}
            }
            + \frac{1}{2}m\omega^2\frac{d}{4\alpha}
        }
        \\
        &=
        \para{
            \frac{\alpha\hbar^2}{2m}
            + \frac{m\omega^2}{8\alpha}
        }dN,
    \end{align}
    which is what we wanted to show.

\section{Computing the drift force and local energy of the full system}
    Computing the gradient of the wavefunction we get
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Bigl[
            \nabla_k
            \Phi_T(\vf{r})
        \Bigr]
        J(\vf{r})
        + \Phi_T(\vf{r})
        \nabla_k J(\vf{r}).
    \end{align}
    The gradient of the Slater permament for particle $k$ is given by
    \begin{align}
        \nabla_k
        \Phi_T(\vf{r})
        &=
        \nabla_k\phi_k
        \prod_{i \neq k}^N\phi_i
        = \frac{\nabla_k\phi_k}{\phi_k}
        \Phi_T(\vf{r}).
    \end{align}
    The gradient of the Jastrow factor is given by
    \begin{align}
        \nabla_k J(\vf{r})
        &=
        J(\vf{r})
        \nabla_k\sum_{m < n}^N u_{mn} \\
        &= J(\vf{r})
        \Biggl(
            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
            \sum_{n = k + 1}^N\nabla_k u_{kn}
        \Biggr)
        \\
        &=
        J(\vf{r})
        \sum_{m \neq k}^N\nabla_k u_{km},
    \end{align}
    where the gradient of the interaction term splits the antisymmetric
    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
    into a single sum. This in total yields the gradient
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Biggl(
            \frac{\nabla_k\phi_k}{\phi_k}
            + \sum_{m \neq k}^N
            \nabla_k u_{km}
        \Biggr)
        \Psi_T(\vf{r}).
        \label{eq:gradient_full_wavefunction}
    \end{align}

@@

    \begin{widetext}
        For the Laplacian we switch a little back and forth between the
        two ways of representing the $\nabla_k$-operator. We thus get
        \begin{align}
            \nabla_k^2 u_{km}
            &=
            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \Biggl[
                \nabla_k \frac{1}{r_{km}}
            \Biggr]
            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
            \dpd[]{u_{km}}{r_{km}}
            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
            \dpd[2]{u_{km}}{r_{km}} \\
            &=
            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \dpd[2]{u_{km}}{r_{km}},
        \end{align}
        where $d$ is again the dimensionality of the problem. In total
        we can state an intermediate version of the Laplacian occuring
        in the local energy as
        \begin{align}
            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
            &=
            \frac{\nabla_k^2\phi_k}{\phi_k}
            + 2\frac{\nabla_k\phi_k}{\phi_k}
            \sum_{m \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \dpd[]{u_{km}}{r_{km}}
            + \sum_{m\neq k}^N
            \Biggl(
                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                + \dpd[2]{u_{km}}{r_{km}}
            \Biggr)
            \nonumber \\
            &\qquad
            +
            \sum_{m, n \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
            \dpd[]{u_{km}}{r_{km}}
            \dpd[]{u_{kn}}{r_{kn}}.
        \end{align}
    \end{widetext}
    Moving on to the derivatives of the interaction terms, $u_{km}$, to
    get an explicit expression for the Laplacian.
    \begin{align}
        \dpd[]{u_{km}}{r_{km}}
        &=
        \frac{a}{r_{km}(r_{km} - a)},
        \\
        \dpd[2]{u_{km}}{r_{km}}
        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
    \end{align}
    The local energy and the drift force can now be found by combining these
    expressions. We will not write out the explicit expressions as
    these will be called by separated functions in our programs.

>>>>

\appendix
\section{Computing the exact energy of the non-interacting system}
    Here we show how to arrive at \autoref{eq:exact_energy}. We need to solve
    two integrals consisting of $N$, $d$-dimensional, Gaussians from
    \autoref{eq:trial_simple_gaussian}. We will label the position in the $j$-th
    coordinate for the $i$-th particle by $x_{ij}$. Starting with the
    normalization factor we get
    \begin{align}
        \bra{\Psi_T}\ket{\Psi_T}
        &=
        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2
        \\
        &=
        \prod_{i = 1}^N
        \prod_{j = 1}^d
        \int\dd x_{ij}\exp(-2\alpha x_{ij}^2)
        \\
        &=
        \prod_{i = 1}^N
        \prod_{j = 1}^d
        \sqrt{\frac{\pi}{2\alpha}}
        = \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{dN}.
    \end{align}
    From the Hamiltonian we get two integrals containing $\vf{r}_i^2$ for each
    particle $i$.
    \begin{align}
        \bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}
        &=
        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2\vf{r}_i^2
        \\
        &=
        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)\vf{r}_i^2
        \nonumber \\
        &\qquad
        \times
        \prod_{j \neq i}^N
        \int\dd\vf{r}_j\exp(-2\alpha\vf{r}_j^2)
        \\
        &=
        \sum_{l = 1}^d
        \int
        \dd x_{il}\exp(-2\alpha x_{il}^2)x_{il}^2
        \nonumber \\
        &\qquad
        \times
            \int
            \prod_{k \neq l}^d
            \dd x_{ik}\exp(-2\alpha x_{ik}^2)
        \nonumber \\
        &\qquad
        \times
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d(N - 1)}
        \\
        &=
        \sum_{l = 1}^d
        \frac{1}{4\alpha}\sqrt{\frac{\pi}{2\alpha}}
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d - 1}
        \nonumber\\
        &\qquad
        \times
        \para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{d(N - 1)}
        \\
        &= \frac{d}{4\alpha}\para{
            \sqrt{\frac{\pi}{2\alpha}}
        }^{dN}.
    \end{align}
    We thus have that
    \begin{align}
        \frac{\bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
        &=
        \frac{d}{4\alpha}.
    \end{align}
    Using \autoref{eq:laplacian_simple_gaussian} we can find the exact energy
    expression we are looking for.
    \begin{align}
        E(\alpha)
        &=
        \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
        \\
        &=
        \sum_{i = 1}^N
        \brac{
            -\frac{\hbar^2}{2m}
            \brak{
                -2d\alpha
                + 4\alpha^2\frac{d}{4\alpha}
            }
            + \frac{1}{2}m\omega^2\frac{d}{4\alpha}
        }
        \\
        &=
        \para{
            \frac{\alpha\hbar^2}{2m}
            + \frac{m\omega^2}{8\alpha}
        }dN,
    \end{align}
    which is what we wanted to show.

\section{Computing the drift force and local energy of the interacting system}
    Using the trial wavefunction from \autoref{eq:initial_trial_wavefunction} we
    start by computing the gradient.
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Bigl[
            \nabla_k
            \Phi_T(\vf{r})
        \Bigr]
        J(\vf{r})
        + \Phi_T(\vf{r})
        \nabla_k J(\vf{r}).
    \end{align}
    The gradient of the Slater permanent for particle $k$ is given by
    \begin{align}
        \nabla_k
        \Phi_T(\vf{r})
        &=
        \nabla_k\phi_k
        \prod_{i \neq k}^N\phi_i
        = \frac{\nabla_k\phi_k}{\phi_k}
        \Phi_T(\vf{r}).
    \end{align}
    The gradient of the Jastrow factor is given by
    \begin{align}
        \nabla_k J(\vf{r})
        &=
        J(\vf{r})
        \nabla_k\sum_{m < n}^N u_{mn} \\
        &= J(\vf{r})
        \Biggl(
            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
            \sum_{n = k + 1}^N\nabla_k u_{kn}
        \Biggr)
        \\
        &=
        J(\vf{r})
        \sum_{m \neq k}^N\nabla_k u_{km},
    \end{align}
    where the gradient of the interaction term splits the antisymmetric
    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
    into a single sum. This in total yields the gradient
    \begin{align}
        \nabla_k\Psi_T(\vf{r})
        &=
        \Biggl(
            \frac{\nabla_k\phi_k}{\phi_k}
            + \sum_{m \neq k}^N
            \nabla_k u_{km}
        \Biggr)
        \Psi_T(\vf{r}).
        \label{eq:gradient_full_wavefunction}
    \end{align}

@@

    \begin{widetext}
        For the Laplacian we switch a little back and forth between the
        two ways of representing the $\nabla_k$-operator. We thus get
        \begin{align}
            \nabla_k^2 u_{km}
            &=
            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \Biggl[
                \nabla_k \frac{1}{r_{km}}
            \Biggr]
            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
            \dpd[]{u_{km}}{r_{km}}
            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
            \dpd[2]{u_{km}}{r_{km}} \\
            &=
            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
            + \dpd[2]{u_{km}}{r_{km}},
        \end{align}
        In total we can state an intermediate version of the Laplacian occuring
        in the local energy as
        \begin{align}
            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
            &=
            \frac{\nabla_k^2\phi_k}{\phi_k}
            + 2\frac{\nabla_k\phi_k}{\phi_k}
            \sum_{m \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \dpd[]{u_{km}}{r_{km}}
            + \sum_{m\neq k}^N
            \Biggl(
                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
                + \dpd[2]{u_{km}}{r_{km}}
            \Biggr)
            \nonumber \\
            &\qquad
            +
            \sum_{m, n \neq k}^N
            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
            \dpd[]{u_{km}}{r_{km}}
            \dpd[]{u_{kn}}{r_{kn}}.
        \end{align}
    \end{widetext}
    Finally we compute the derivatives of the interaction terms, $u_{km}$, to
    get an explicit expression for the Laplacian.
    \begin{align}
        \dpd[]{u_{km}}{r_{km}}
        &=
        \frac{a}{r_{km}(r_{km} - a)},
        \\
        \dpd[2]{u_{km}}{r_{km}}
        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
    \end{align}
    The local energy and the drift force can now be found by combining these
    expressions. We will not write out the explicit expressions as
    these will be called by separated functions in our programs.
